### 2019-06-10 R/Analyse changes in pelagic stocks.Rmd
Updates on several codes

# -----------------------------------------------------------------------------------------------
# Analyse changes in pelagic stocks.Rmd
#
# 07/01/2019 first coding; adapted from ICES advice and assessment summaries.Rmd
# 10/01/2018 made small changes to the table
# -----------------------------------------------------------------------------------------------

@@

# create subset for advice
d <-
  iAdvice %>%

  filter(grepl("her-47|mac-67|mac-nea|whb-c|whb-n|hom-c|hom-w|her-3a22|her-noss", 
               stockkeylabelold)) %>%
  filter(!grepl("benchmark", purpose)) %>% 
  filter(assessmentyear >= 2008) %>% 
  ungroup()

# create subset for assessments
e <-
  iAssess %>% 
  filter(stockkeylabelold %in% unique(d$stockkeylabelold)) %>% 
  filter(assessmentyear == 2018) %>% 
  filter(year >= 2008) %>% 
  filter(year <= assessmentyear) %>% 
  filter(adviceonstock == TRUE)

@@

d %>% 
  dplyr::select(stockkeylabelold, assessmentyear, fpa, fmsy) %>% 
  gather(key=var, value=value, fpa:fmsy) %>%

@@

d %>% 
  dplyr::select(stockkeylabelold, assessmentyear, blim, msybtrigger) %>% 
  gather(key=var, value=value, blim:msybtrigger) %>%

@@

d %>% 
  dplyr::select(stockkeylabelold, tacyear, adviceonstock, advisedlandingsmax) %>% 
  ggplot(aes(x=tacyear)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

  geom_point(aes(y       = advisedlandingsmax, 
                 colour  = ifelse(adviceonstock==TRUE,'advice','replaced'))) +
  geom_line(aes(y        = advisedlandingsmax, 
                colour   = ifelse(adviceonstock==TRUE,'advice','replaced'),
                linetype = ifelse(adviceonstock==TRUE,'advice',"replaced") )) +

  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=0) + 
  labs(y="Advice Landings Max (tonnes)") +
  scale_color_manual(values = c("red", "blue")) +
  scale_linetype_manual(values = c("solid", "blank")) +
  facet_wrap(~stockkeylabelold, scales="free_y")

@@

d %>% 
  dplyr::select(stockkeylabelold, assessmentyear, adviceonstock, ssbay) %>% 
  rename(year = assessmentyear) %>%

@@

  geom_point(aes(y=ssbay,
                 colour   = ifelse(adviceonstock==TRUE,'advice','replaced'))) +
  geom_line(data=e,
            aes(y = stocksize )) +

@@

@@

d %>% 
  dplyr::select(stock=stockkeylabelold, assessyear=assessmentyear, purpose, fpa, fmsy, blim,
                msybtrig=msybtrigger, ssb_stf=ssbay, tacyear, fadv=fadvmax,
                advlandings=advisedlandingsmax)  %>% 
  mutate(msybtrig = comma_format(big.mark ="") (msybtrig)) %>% 
  arrange(stock, assessyear) %>% 
  pandoc.table(., 
             style        = "simple",
             split.tables = 180, 
             justify      = "right",
             missing      =".",
             big.mark     = '', 
             round        = c(0,0,0,2,2,0,0,0,0,2,0,0,0,0,0,0,0,0,0,0) )

@@

d %>% 
  dplyr::select(stock=stockkeylabelold, assessyear=assessmentyear, purpose, fpa, fmsy, blim,
                msybtrig=msybtrigger, ssb_stf=ssbay, tacyear, fadv=fadvmax,
                advlandings=advisedlandingsmax)  %>% 
  mutate(msybtrig = comma_format(big.mark ="") (msybtrig)) %>% 
  arrange(stock, assessyear) %>% 
  write.csv(., file="advice.csv")

>>>>

# -----------------------------------------------------------------------------------------------
# Analyse changes in pelagic stocks.Rmd
#
# 07/01/2019 first coding; adapted from ICES advice and assessment summaries.Rmd
# 10/01/2018 made small changes to the table
# 04/05/2019 updated for WKRRMAC 2019; focus on mackerel
# 06/06/2019 changed the data frame d; now also has purpose
# -----------------------------------------------------------------------------------------------

@@

# create subset for advice
adv <-
  iAdvice %>%

  filter(grepl("her-47|mac-67|mac-nea|whb-c|whb-n|hom-c|hom-w|her-3a22|her-noss",
               stockkeylabelold)) %>%
  # filter(grepl("mac-67|mac-nea", 
  #              stockkeylabelold)) %>%

  mutate(purpose = ifelse(grepl("bench", purpose), "benchmark", purpose)) %>% 
  mutate(purpose = ifelse(grepl("initial", purpose), "replaced", purpose)) %>%

  filter(purpose %in% c("advice","replaced", "benchmark")) %>% 
  mutate(id      = paste0(stockkeylabelold, assessmentyear, assessmentdate, purpose)) %>%

  filter(assessmentyear >= 2008) %>% 
  ungroup()

adv_ssb <-
  adv %>% 
  dplyr::select(stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew, 
                id, assessmentyear, purpose, ssbay) %>% 
  rename(stocksize = ssbay) %>% 
  mutate(year = assessmentyear)

# adv %>% filter(stockkeylabelold =="mac-nea", assessmentyear >= 2015) %>% View()
# adv %>% distinct(purpose) %>% View()
# create subset for assessments

ass <-
  iAssess %>% 
  filter(stockkeylabelold %in% unique(adv$stockkeylabelold)) %>% 
  filter(assessmentyear %in% unique(adv$assessmentyear)) %>% 
  filter(year <  assessmentyear) %>% 
  filter(year >= 2000) %>%

  group_by(stockkeylabelold, assessmentyear) %>% 
  mutate(n       = n_distinct(paste0(stockkeylabelold, assessmentyear, adviceonstock))) %>%

  mutate(purpose = ifelse(grepl("bench", purpose), "benchmark", purpose)) %>% 
  mutate(purpose = ifelse(grepl("initial", purpose), "replaced", purpose)) %>%

  mutate(id      = paste0(stockkeylabelold, assessmentyear, assessmentdate, purpose)) %>% 
  mutate(ayear   = substr(as.character(assessmentyear), 3,4)) %>% 
  mutate(source  = "assessment") %>% 
  arrange(assessmentyear, assessmentdate, purpose, year) %>% 
  ungroup()

ass_final <-
  ass %>% 
  group_by(stockkeylabelold) %>%
  filter(purpose == "advice") %>% 
  filter(assessmentyear == max(assessmentyear, na.rm=TRUE)) %>% 
  ungroup()

ass_5year <-
  ass %>% 
  group_by(assessmentyear, purpose) %>% 
  filter(year >= (max(year, na.rm=TRUE)-4) ) %>%

  # add the current year estimate from adv
  bind_rows(adv_ssb) %>% 
  arrange(assessmentyear, assessmentdate, purpose, year) %>% 
  ungroup()

ass_10year <-
  ass %>% 
  group_by(assessmentyear, purpose) %>% 
  filter(year >= (max(year, na.rm=TRUE)-9) ) %>%

  # add the current year estimate from adv
  bind_rows(adv_ssb) %>% 
  arrange(assessmentyear, assessmentdate, purpose, year) %>% 
  ungroup()

# ass <-
#   iAssess %>% 
#   group_by(assessmentyear) %>% 
#   filter(stockkeylabelold %in% unique(adv$stockkeylabelold)) %>% 
#   filter(assessmentyear == max(assessmentyear, na.rm=TRUE)) %>% 
#   filter(year <= assessmentyear) %>% 
#   
#   # count number of assessments per year; keep assessment if n=1 or otherwise if adviceonstock is TRUE
#   group_by(stockkeylabelold, assessmentyear) %>% 
#   mutate(n = n_distinct(paste0(stockkeylabelold, assessmentyear, adviceonstock))) %>% 
#   filter(n == 1 | (assessmentdate == max(assessmentdate, na.rm=TRUE)) ) %>% 
#   filter(year >= 2008) %>% 
#   ungroup()
# #   filter(n == 1 | (assessmentdate == max(assessmentdate, na.rm=TRUE)) )

# adv %>%
#   ungroup() %>%
#   distinct(stockkeylabelold, assessmentyear,  benchmark, adviceonstock, n, .keep_all = FALSE) %>%
#   arrange(stockkeylabelold, assessmentyear, benchmark, adviceonstock, n) %>%
#   View()

@@

adv %>% 
  filter(purpose == "advice") %>%

  dplyr::select(stockkeylabelold, assessmentyear, fpa, fmsy) %>% 
  gather(key=var, value=value, fpa:fmsy) %>% 
  # View()

@@

adv %>% 
  filter(purpose == "advice") %>%

  dplyr::select(stockkeylabelold, assessmentyear, blim, msybtrigger) %>% 
  gather(key=var, value=value, blim:msybtrigger) %>%

@@

adv %>%

  filter(purpose %in% c("advice","replaced")) %>%

  dplyr::select(stockkeylabelold, tacyear, purpose, advisedlandingsmax) %>%

  ggplot(aes(x=tacyear)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

  geom_point(aes(y       = advisedlandingsmax, colour=purpose)) +
  geom_line(aes(y        = advisedlandingsmax, 
                colour   = purpose,
                linetype = purpose )) +

  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=0) + 
  labs(y="Advice Landings Max (tonnes)") +
  scale_color_manual(values = c("red", "blue")) +
  scale_linetype_manual(values = c("solid", "blank", "dotted")) +
  facet_wrap(~stockkeylabelold, scales="free_y")

@@

# subset of ass
t <-
  ass %>% 
  filter(purpose == "advice") %>% 
  group_by(stockkeylabelold) %>% 
  filter(assessmentyear == max(assessmentyear)  )

adv %>%

  filter(purpose %in% c("advice","replaced")) %>%

  dplyr::select(stockkeylabelold, assessmentyear, purpose, ssbay) %>% 
  rename(year = assessmentyear) %>% 
  # View()

@@

  geom_point(aes(y=ssbay,
                 colour   = purpose)) +
  geom_line(data=t,
            aes(y = stocksize )) +

@@

# NOT SURE WHAT THIS IS ADDING; NOT SHOWN

# now based on assessment
t <-
  ass %>% 
  filter(year == assessmentyear - 1 ) %>% 
  mutate(colour  = paste0(adviceonstock,benchmark),
         purpose = ifelse(grepl("bench", purpose), "benchmark", purpose),
         purpose = ifelse(grepl("initial", purpose), "replaced", purpose))

# glimpse(adv)

ass %>% 
  group_by(stockkeylabelold) %>% 
  filter(assessmentyear == max(assessmentyear, na.rm=TRUE)) %>% 
  dplyr::select(stockkeylabelold, assessmentyear, year, adviceonstock, stocksize) %>% 
  # rename(year = assessmentyear) %>% 
  # View()

  # geom_point(aes(y=ssbay,
  #                colour   = ifelse(adviceonstock==TRUE,'advice','replaced'))) +
  geom_line(aes(y = stocksize )) +
  geom_point(aes(y = stocksize )) +

  geom_point(data=t, aes(x=year, y=stocksize, colour=purpose)) +
  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=0) + 
  labs(y="SSB") +
  # scale_color_manual(values = c("red", "blue")) +
  facet_wrap(~stockkeylabelold, scales="free_y")
```

```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

# Historical retro with all the variants

ass_5year %>%

  group_by(stockkeylabelold, assessmentyear) %>% 
  filter(year < max(assessmentyear, na.rm=TRUE) ) %>%

  ggplot(aes(x=year, y=stocksize, group=id, colour=purpose)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

  # geom_point(data=adv_ssb, aes(colour = purpose)) +

  geom_line(aes(size=purpose)) +
  geom_line(data=ass_final, aes(colour=purpose ), size=1.5) +
  # geom_dl(aes(label  = ayear, group=purpose), method = list(dl.combine("last.points"), cex = 0.8)) +

  scale_colour_manual(values=c(advice = "black",benchmark = "blue",replaced = "red")) +
  scale_size_manual  (values=c(advice = 0.6, benchmark = 0.9, replaced = 0.9)) +

  scale_x_continuous(breaks=pretty_breaks()) +

  expand_limits(y=0) + 
  labs(y="SSB",y="") +
  facet_wrap(~stockkeylabelold, scales="free_y")

# distinct(ass_5year, purpose) %>% View()
# ass %>% filter(grepl("bench", purpose)) %>% View()

@@

adv %>% 
  dplyr::select(stock=stockkeylabelold, assessyear=assessmentyear, purpose, fpa, fmsy, blim,
                msybtrig=msybtrigger, ssb_stf=ssbay, tacyear, fadv=fadvmax,
                advlandings=advisedlandingsmax)  %>% 
  mutate(msybtrig = comma_format(big.mark ="") (msybtrig)) %>% 
  arrange(stock, assessyear) %>% 
  pandoc.table(., 
             style        = "simple",
             split.tables = 180, 
             justify      = "right",
             missing      =".",
             big.mark     = '', 
             round        = c(0,0,0,2,2,0,0,0,0,2,0,0,0,0,0,0,0,0,0,0) )

@@

adv %>% 
  dplyr::select(stock=stockkeylabelold, assessyear=assessmentyear, purpose, fpa, fmsy, blim,
                msybtrig=msybtrigger, ssb_stf=ssbay, tacyear, fadv=fadvmax,
                advlandings=advisedlandingsmax)  %>% 
  mutate(msybtrig = comma_format(big.mark ="") (msybtrig)) %>% 
  arrange(stock, assessyear) %>% 
  write.csv(., file="advice.csv")
